---
title: "BACS HW (Week 15)"
author: '108020024'
date: "due on 05/28 (Sun) helped by 108020033"
output:
  html_document:
    df_print: default
  pdf_document:
    latex_engine: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE) 

```

```{r}
#install.packages("seminr")
library(seminr)

```


### Question 1)  Composite Path Models using PLS-PM


### a) Create a PLS path model using SEMinR, with all the following characteristics:


```{r}

sq <- read.csv("security_data_sem.csv")
```

### i) Measurement model – all constructs are measured as composites:

```{r}
sec_mm <- constructs(
composite("REP", multi_items("PREP", 1:4)),
composite("INV", multi_items("PINV", 1:3)),
composite("SEC", multi_items("PSEC", 1:4)),
composite("POL", multi_items("PPSS", 1:3)),
composite("TRUST", multi_items("TRST", 1:4)),
composite("FAML", single_item("FAML1")),
interaction_term(iv="REP", moderator="POL", method=orthogonal)
)
```

### ii) Structural Model – paths between constructs as shown in this causal model:

```{r}

sec_sm <- relationships(
paths(from = c("REP", "INV", "POL" , "FAML" , "REP*POL"), to = "SEC"),
paths(from = "SEC", to = "TRUST")
)

```


### b) Show us the following results in table or figure formats:


### i) Plot a figure of the estimated model

```{r}
sec_pls <- estimate_pls(data = sq,
  measurement_model = sec_mm,
  structural_model = sec_sm)
plot(sec_pls)
```


### ii) Weights and loadings of composites


```{r}
summary(sec_pls)$weight
summary(sec_pls)$composite_scores
```

### iii) Regression coefficients of paths between factors

```{r}
summary(sec_pls)$paths

```


### iv) Bootstrapped path coefficients: t-values, 95% CI


```{r}
boot_pls <- bootstrap_model(sec_pls, nboot = 1000)
summary(boot_pls)
```


### Question 2) Common-Factor Models using CB-SEM


### a) Create a common factor model using SEMinR, with the following characteristics:

```{r}
sec_mm <- constructs(
composite("REP", multi_items("PREP", 1:4)),
composite("INV", multi_items("PINV", 1:3)),
composite("SEC", multi_items("PSEC", 1:4)),
composite("POL", multi_items("PPSS", 1:3)),
composite("TRUST", multi_items("TRST", 1:4)),
composite("FAML", single_item("FAML1")),
interaction_term(iv="REP", moderator="POL", method=orthogonal)
)
sec_mm_reflective = as.reflective(sec_mm)
```

```{r}

sec_sm <- relationships(
paths(from = c("REP", "INV", "POL" , "FAML" , "REP*POL"), to = "SEC"),
paths(from = "SEC", to = "TRUST")
)

```


### b) Show us the following results in table or figure formats


### i) Plot a figure of the estimated model

```{r}
sec_pls_reflective <- estimate_cbsem(data = sq,
  measurement_model = sec_mm_reflective,
  structural_model = sec_sm)
plot(sec_pls_reflective)
```


### ii) loadings of composites


```{r}
sec_pls_rfsum<-summary(sec_pls_reflective)
sec_pls_rfsum$loadings
```

### iii) Regression coefficients of paths between factors, and their p-values


```{r}
sec_pls_rfsum$paths

```




