---
title: "BACS HW (Week 1)"
author: "108020024"
date: due on 02/19 (Sun)
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Before working on HW1, first complete the tutorial of Swirl assigned this week to get familial with r programming language. 


### 1.What is the 5th element in the original list of ages?

For problem 1, first read the text file into a data frame named "customer".
```{r}
customers <- read.table("customers.txt",head=TRUE) #read the target txt file
customers[5,] #print the 5th element
```
Since the data frame has only one variable "age", just print the fifth row of customer,and the 5th element in the original list of ages is 45.

### 2. What is the fifth lowest age?

For problem 2, sort the data into order.
```{r}
customers_sorted <- sort(customers$age) #sort the age list from smallest to largest
customers_sorted[5] #get the fifth lowest age
```
So 19 is the fifth lowest age.


### 3. Extract the five lowest ages together.

```{r}
customers_sorted[1:5] #get the 1 to 5 elements.
```
The five lowest ages are [18 19 19 19 19].


### 4. Get the five highest ages by first sorting them in decreasing order first.

```{r}
customers_sorted2 <- sort(customers$age,decreasing = TRUE) #sort from largest to smallest
customers_sorted2[1:5] #get the 1 to 5 elements.
```
The five highest ages are [85 83 82 82 81].

### 5. What is the average (mean) age?

```{r}
age_mean <- mean(customers$age) #calculate the mean of age
age_mean
```
The average age is 46.80702.


### 6. What is the standard deviation of ages?

```{r}
age_sd <- sd(customers$age) #calculate the standard deviation of age
age_sd
```
The standard deviation of ages is 16.3698.

### 7. Make a new variable called age_diff, with the difference between each age and the mean age.

```{r}
age_diff <- c(customers$age - age_mean)
```

### 8. What is the average “difference between each age and the mean age”?

```{r}
mean(age_diff) #calculate the mean of age_diff
```

The average “difference between each age and the mean age” is -1.623275e-15, however the computer is trying to say 0. The really small number is because of the imprecision that computers makes doing calculation.


### 9. Visualize the raw data as we did in class: (a) histogram, (b) density plot, (c) boxplot+stripchart

### (a) histogram
```{r}
hist(customers$age, main = "Histogram of Age", xlab = "Age")
```

### (b) density plot
```{r}
plot(density(customers$age), main = "Histogram of Age", xlab = "Age")
```

### (c) boxplot+stripchart
```{r}

boxplot(customers$age, main = "Boxplot+Stripchart of Age")
stripchart(customers$age, add = TRUE, vertical = TRUE,
           method = "jitter", col = "red", pch = 16)
```